Revisiting the Relationship between Short-Chain Fatty Acids, the Microbiota, and Colorectal Tumors

### 2018-08-06 submission/manuscript.Rmd
3rd paragraph results - ada suggestions

**Total significant positive correlations between OTU relative abundance and SCFA concentration were similar for individuals without tumors and patients with adenomas or carcinomas.** Having found no difference between individuals without tumors and patients with adenomas or carcinomas in SCFA concentrations or genes coding for enzymes involved with SCFA synthesis, we next investigated if specific OTUs correlated with SCFA concentrations. Using Spearmans rho we found that the majority of significant OTU correlations were to taxa from *Clostridiales*, *Lachnospiraceae*, and *Ruminococcaceae* [Figure 3 & Table S5]. A similar pattern was observed when using high/low SCFA groups based on the overall median concentration for that specific SCFA [Figure S1 & Table S6]. There was a noticeably higher number of significant negative correlations associated with patients with adenomas for all SCFAs tested [Figure 3]. In particular, OTUs from the *Ruminococcaceae* family had the largest share of these negative correlations within patients with adenomas  [Figure 3]. Although patients with adenomas had more positive correlations between OTUs and SCFA concentrations, their total number was more similar to individuals without tumors or patients with carcinomas versus the analogous comparison for the number of negative correlations [Figure 3]. The number of positive correlations between OTUs and SCFA concentrations was similar between individuals without tumors and patients with a carcinoma [Figure 3]. Overall, these results suggest that the resident taxa that may change the most due to colon tumors may not be ones that are responsible for the production of acetate, butyrate, or propionate.

>>>>

**The number OTUs positively associated with SCFA concentration were similar between individuals without tumors and patients with adenomas or carcinomas.** Having found no difference between individuals without tumors and patients with adenomas or carcinomas in SCFA concentrations or genes encoding enzymes involved with SCFA synthesis, we next investigated if specific OTUs correlated with SCFA concentrations. The main goal of this analysis was to identify if there were OTUs that were significantly associated with SCFA concentrations and if this was different between individuals that did not have tumors, had an adenoma, or had a carcinoma. To accomplish this we used Spearmans rho, a non-parametric measure of association, and tested if there was a correlation that was signficantly greater than zero. We found that taxa from *Clostridiales*, *Lachnospiraceae*, and *Ruminococcaceae* dominated statistically significant OTU correlations [Figure 3 & Table S5]. There was a noticeably higher number of significant negative correlations associated with patients with adenomas for all SCFAs tested [Figure 3]. In particular, OTUs from the *Ruminococcaceae* family had the largest share of these negative correlations within patients with adenomas [Figure 3]. Patients with adenomas also had more positive correlations between OTUs and SCFA concentrations, but their total number was more similar to individuals without tumors or patients with carcinomas versus the analogous comparison for the number of negative correlations [Figure 3]. Additionally, the number of positive correlations between OTUs and SCFA concentrations was similar between individuals without tumors and patients with a carcinoma [Figure 3]. Finally, when we used high/low SCFA groups based on the overall median concentration for each SCFA instead of SCFA concentrations a similar pattern was still observed [Figure S1 & Table S6]. Overall, these results suggest that the resident taxa that may change the most due to colon tumors may not be ones that are responsible for the production of acetate, butyrate, or propionate.

### 2018-08-06 submission/manuscript.tex
3rd paragraph results - ada suggestions

\textbf{Total significant positive correlations between OTU relative
abundance and SCFA concentration were similar for individuals without
tumors and patients with adenomas or carcinomas.} Having found no
difference between individuals without tumors and patients with adenomas
or carcinomas in SCFA concentrations or genes coding for enzymes
involved with SCFA synthesis, we next investigated if specific OTUs
correlated with SCFA concentrations. Using Spearmans rho we found that
the majority of significant OTU correlations were to taxa from
\emph{Clostridiales}, \emph{Lachnospiraceae}, and \emph{Ruminococcaceae}
{[}Figure 3 \& Table S5{]}. A similar pattern was observed when using
high/low SCFA groups based on the overall median concentration for that
specific SCFA {[}Figure S1 \& Table S6{]}. There was a noticeably higher
number of significant negative correlations associated with patients
with adenomas for all SCFAs tested {[}Figure 3{]}. In particular, OTUs
from the \emph{Ruminococcaceae} family had the largest share of these
negative correlations within patients with adenomas {[}Figure 3{]}.
Although patients with adenomas had more positive correlations between
OTUs and SCFA concentrations, their total number was more similar to
individuals without tumors or patients with carcinomas versus the
analogous comparison for the number of negative correlations {[}Figure
3{]}. The number of positive correlations between OTUs and SCFA
concentrations was similar between individuals without tumors and
patients with a carcinoma {[}Figure 3{]}. Overall, these results suggest
that the resident taxa that may change the most due to colon tumors may
not be ones that are responsible for the production of acetate,
butyrate, or propionate.

>>>>

\textbf{The number OTUs positively associated with SCFA concentration
were similar between individuals without tumors and patients with
adenomas or carcinomas.} Having found no difference between individuals
without tumors and patients with adenomas or carcinomas in SCFA
concentrations or genes encoding enzymes involved with SCFA synthesis,
we next investigated if specific OTUs correlated with SCFA
concentrations. The main goal of this analysis was to identify if there
were OTUs that were significantly associated with SCFA concentrations
and if this was different between individuals that did not have tumors,
had an adenoma, or had a carcinoma. To accomplish this we used Spearmans
rho, a non-parametric measure of association, and tested if there was a
correlation that was signficantly greater than zero. We found that taxa
from \emph{Clostridiales}, \emph{Lachnospiraceae}, and
\emph{Ruminococcaceae} dominated statistically significant OTU
correlations {[}Figure 3 \& Table S5{]}. There was a noticeably higher
number of significant negative correlations associated with patients
with adenomas for all SCFAs tested {[}Figure 3{]}. In particular, OTUs
from the \emph{Ruminococcaceae} family had the largest share of these
negative correlations within patients with adenomas {[}Figure 3{]}.
Patients with adenomas also had more positive correlations between OTUs
and SCFA concentrations, but their total number was more similar to
individuals without tumors or patients with carcinomas versus the
analogous comparison for the number of negative correlations {[}Figure
3{]}. Additionally, the number of positive correlations between OTUs and
SCFA concentrations was similar between individuals without tumors and
patients with a carcinoma {[}Figure 3{]}. Finally, when we used high/low
SCFA groups based on the overall median concentration for each SCFA
instead of SCFA concentrations a similar pattern was still observed
{[}Figure S1 \& Table S6{]}. Overall, these results suggest that the
resident taxa that may change the most due to colon tumors may not be
ones that are responsible for the production of acetate, butyrate, or
propionate.
